---
title: "MSIA 401 HW1"
author: "Yijun Wu"
date: "9/21/2020"
output: html_document
---

#### 2.1

Set $\frac{\partial Q}{\partial \beta}=0$ to minimize Q and then get the expression for $\beta$:
$$\begin{equation}
  \begin{split}
  Q 
  &= \sum^n_{i=1}(y_i-\hat\beta x_i)^2 \\
  \frac{\partial Q}{\partial \beta}
  &=  \sum^n_{i=1}2(-x_i)(y_i-\hat\beta x_i) =0\\
  0
  &= -2\sum^n_{i=1}(x_iy_i-\hat\beta {x_i}^2) \\
  \hat\beta \sum^n_{i=1} {x_i}^2 
  &= \sum^n_{i=1}x_iy_i \\
  \hat\beta
  &= \frac{\sum^n_{i=1}x_iy_i}{\sum^n_{i=1} {x_i}^2 }
  \end{split}
\end{equation}$$

#### 2.3
Set $\frac{\partial Q}{\partial \beta}=0$ to minimize Q and then get the expression for $\beta$:
$$\begin{equation}
  \begin{split}
  Q 
  &= \sum^n_{i=1}w_i(y_i-\hat\beta x_i)^2 \\
  Q
  &= \sum^n_{i=1}w_i(y_i^2-2\hat\beta x_i y_i+{\hat\beta}^2{x_i}^2)\\
  \frac{\partial Q}{\partial \beta}
  &=  \sum^n_{i=1}w_i(-2x_i y_i+2\hat\beta{x_i}^2) =0\\
  \hat\beta \sum^n_{i=1} w_i {x_i}^2
  &= \sum^n_{i=1}w_i x_i y_i \\
  \hat\beta
  &= \frac{\sum^n_{i=1}w_i x_iy_i}{\sum^n_{i=1} w_i {x_i}^2 }
  \end{split}
\end{equation}$$


#### 2.8
```{r}
father_s <- 64
father_t <- 72
sx <- 2.7
sy <- 2.7
father_bar <- 68
son_bar <- 69
r1 <- 0.25
r2 <- 0.75
```

```{r}
# when r = 0.25
son_s_1 <- r1*(sx/sy)*(father_s-father_bar)+son_bar 
son_t_1 <- r1*(sx/sy)*(father_t-father_bar)+son_bar 
#when r = 0.75
son_s_2 <- r2*(sx/sy)*(father_s-father_bar)+son_bar 
son_t_2 <- r2*(sx/sy)*(father_t-father_bar)+son_bar 
```

```{r}
#result
table <- data.frame("Father shorter"= c(son_s_1, son_s_2), "Father taller" = c(son_t_1, son_t_2))
row.names(table) <- c("r=.25", "r=.75")
table
```

#### 2.9

```{r}
RR_data <- read.csv("IBM-Apple-SP500 RR Data.csv", header = TRUE, skip = 1)
RR_data$S.P.500 <- as.numeric(gsub("\\%", "", RR_data$S.P.500 ))
RR_data$IBM <- as.numeric(gsub("\\%", "", RR_data$IBM ))
RR_data$Apple <- as.numeric(gsub("\\%", "", RR_data$Apple ))
```

##### a.
```{r}
par(mfrow = c(1,2))
plot(RR_data$S.P.500, RR_data$IBM, xlab = "Rate of Return of S&P 500", ylab = "Rate of Return on IBM", main = "IBM vs S&P 500", ylim = c(-35, 30))
plot(RR_data$S.P.500, RR_data$Apple, xlab = "Rate of Return of S&P 500", ylab = "Rate of Return on Apple", main = "Applevs S&P 500" , ylim = c(-35, 30))
```

##### b.
```{r}
apple_sp500 <- lm(Apple ~ S.P.500, data = RR_data)
ibm_sp500 <- lm(IBM ~ S.P.500, data = RR_data)
ibm_beta <- as.numeric(coef(ibm_sp500)[2])
apple_beta <- as.numeric(coef(apple_sp500)[2])
c("Beta of IBM" = ibm_beta, "Beta of Apple" = apple_beta)
```


##### c.
```{r}
sd_sp500 <- sd(RR_data$S.P.500)
sd_ibm <- sd(RR_data$IBM)
sd_apple <- sd(RR_data$Apple)
r_ibm <- cor(RR_data$S.P.500, RR_data$IBM)
r_apple <- cor(RR_data$S.P.500, RR_data$Apple)
cor(RR_data[2:4])
ibm_beta_c <- r_ibm*sd_ibm/sd_sp500
apple_beta_c <- r_apple*sd_apple/sd_sp500
c("Beta of IBM" = ibm_beta_c, "Beta of Apple" = apple_beta_c)
```




#### 2.10
```{r}
steakprices <- read.csv("steakprices.CSV")
steakprices$Chuck.Price <- as.numeric(gsub("\\$", "", steakprices$Chuck.Price ))
steakprices$PortHse.Price <- as.numeric(gsub("\\$", "", steakprices$PortHse.Price ))
steakprices$RibEye.Price <- as.numeric(gsub("\\$", "", steakprices$RibEye.Price ))
```

##### a.
```{r}
chuck <- lm(log(Chuck.Qty) ~ log(Chuck.Price), data = steakprices)
chuck_pe <- as.numeric(coef(chuck)[2])
porthse_pe <- as.numeric(coef(lm(log(PortHse.Qty) ~ log(PortHse.Price), data = steakprices))[2])
ribeye_pe <- as.numeric(coef(lm(log(RibEye.Qty) ~ log(RibEye.Price), data = steakprices))[2])
pe = c("Chuck" = chuck_pe, "Porter House" = porthse_pe, "Rib Eye" = ribeye_pe)
pe
```
Yes, their price elasticities are in the expexted order, with the most expensive Porter House having the greatest price elasticity and the cheapest Chuck has the smallest price elasticity.

##### b.
```{r}
(1-1.1^(pe))*100
```
With 10% increase in price, the demand of Chuck will decrease by 12.23%, the demand of Porter House will decrease by 22.37%, and the demand of Rib Eye will decrease by 12.87%. 


#### 2.11
```{r}
cancer <- read.csv("Cancer Data.csv")
```

##### a.
```{r}
par(mfrow = c(2,2))
plot(cancer$CIG.SMOKED, cancer$BLADDER, xlab = "Cigarettes Smoked", ylab = "Number of Deatths due to Bladder Cancer")
plot(cancer$CIG.SMOKED, cancer$LUNG, xlab = "Cigarettes Smoked", ylab = "Number of Deatths due to Lung Cancer")
plot(cancer$CIG.SMOKED, cancer$KIDNEY, xlab = "Cigarettes Smoked", ylab = "Number of Deatths due to Kidney Cancer")
plot(cancer$CIG.SMOKED, cancer$LEUKEMIA, xlab = "Cigarettes Smoked", ylab = "Number of Deatths due to Leukemia")
```

##### b.
```{r}
#Correlation Matrix
cor(cancer[-1])
r <- cor(cancer[-1])[1, -1]
#t-test statistics
t_stat = r*sqrt(nrow(cancer)-2)/sqrt(1-r^2)
t_stat
#p(>|t|)
pt(t_stat, nrow(cancer)-2, lower.tail = FALSE)*2
```
Based on the t-test, bladder cancer, lung cancer and kidney cancer are all significantly correlated with cigarette smoking. Leukemia is not significantly correlated with cigarette smoking. 
